---
title: "mlr"
author: "Nguyễn Ngọc Bình"
date: "November 19, 2019"
output:
  html_document: 
    code_download: true
    code_folding: hide
    number_sections: yes
    theme: "default"
    toc: TRUE
    toc_float: TRUE
    dev: 'svg'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
    
```
# Getting Data

```{r}
#load libraries and data
install.packages("mlr")
library(mlr)
train <- read.csv("train_loan.csv", na.strings = c(""," ",NA))
test <- read.csv("test_Y3wMUE5.csv", na.strings = c(""," ",NA))


summarizeColumns(train)
```

# Exploring Data
```{r}
summarizeColumns(train)
summarizeColumns(test)

```

```{r}
hist(train$ApplicantIncome, breaks = 300, main = "Applicant Income Chart",xlab = "ApplicantIncome")
hist(train$CoapplicantIncome, breaks = 100,main = "Coapplicant Income Chart",xlab = "CoapplicantIncome")
```

```{r}
boxplot(train$ApplicantIncome)
```

Let’s change the class of Credit_History to factor. Remember, the class factor is always used for categorical variables.

```{r}
train$Credit_History <- as.factor(train$Credit_History)
test$Credit_History <- as.factor(test$Credit_History)
```

```{r}
class(train$Credit_History)
```

You can further scrutinize the data using:

```{r}
summary(train)
summary(test)
```

We find that the variable Dependents has a level 3+ which shall be treated too. It’s quite simple to modify the name levels in a factor variable. It can be done as:

```{r}
#rename level of Dependents
levels(train$Dependents)[4] <- "3"
levels(test$Dependents)[4] <- "3"
```

# Missing Value Imputation
```{r}
#impute missing values by mean and mode
imp <- impute(train, classes = list(factor = imputeMode(), integer = imputeMean()), dummy.classes = c("integer","factor"), dummy.type = "numeric")
imp1 <- impute(test, classes = list(factor = imputeMode(), integer = imputeMean()), dummy.classes = c("integer","factor"), dummy.type = "numeric")
```

$data attribute of imp function contains the imputed data.

```{r}
imp_train <- imp$data
imp_test <- imp1$data
```

Now, we have the complete data. You can check the new variables using:

```{r}
summarizeColumns(imp_train)
summarizeColumns(imp_test)
```


```{r}
listLearners("classif", check.packages = TRUE, properties = "missings")[c("class","package")]
```

Did you notice a disparity among both data sets? No ? See again. The answer is Married.dummy variable exists only in imp_train and not in imp_test. Therefore, we’ll have to remove it before modeling stage.

Optional: You might be excited or curious to try out imputing missing values using a ML algorithm. In fact, there are some algorithms which don’t require you to impute missing values. You can simply supply them missing data. They take care of missing values on their own. Let’s see which algorithms are they:

```{r}
listLearners("classif", check.packages = TRUE, properties = "missings")[c("class","package")]
```

However, it is always advisable to treat missing values separately. Let’s see how can you treat missing value using rpart:

```{r}
rpart_imp <- impute(train, target = "Loan_Status",
                    classes = list(numeric = imputeLearner(makeLearner("regr.rpart")),
                                   factor = imputeLearner(makeLearner("classif.rpart"))),
                    dummy.classes = c("numeric","factor"),
                    dummy.type = "numeric")
```

# Feature Engineering

Feature Engineering is the most interesting part of predictive modeling. So, feature engineering has two aspects: Feature Transformation and Feature Creation. We’ll try to work on both the aspects here.

At first, let’s remove outliers from variables like ApplicantIncome, CoapplicantIncome, LoanAmount. There are many techniques to remove outliers. Here, we’ll cap all the large values in these variables and set them to a threshold value as shown below:

```{r}
# for train data set
cd <- capLargeValues(imp_train, target = "Loan_Status",cols = c("ApplicantIncome"),threshold = 40000)
cd <- capLargeValues(cd, target = "Loan_Status",cols = c("CoapplicantIncome"),threshold = 21000)
cd <- capLargeValues(cd, target = "Loan_Status",cols = c("LoanAmount"),threshold = 520)

#rename the train data as cd_train
cd_train <- cd

#add a dummy Loan_Status column in test data
imp_test$Loan_Status <- sample(0:1,size = 367,replace = T)

cde <- capLargeValues(imp_test, target = "Loan_Status",cols = c("ApplicantIncome"),threshold = 33000)
cde <- capLargeValues(cde, target = "Loan_Status",cols = c("CoapplicantIncome"),threshold = 16000)
cde <- capLargeValues(cde, target = "Loan_Status",cols = c("LoanAmount"),threshold = 470)

#renaming test data
cd_test <- cde
```

I’ve chosen the threshold value with my discretion, after analyzing the variable distribution. To check the effects, you can do summary(cd_train$ApplicantIncome) and see that the maximum value is capped at 33000.

In both data sets, we see that all dummy variables are numeric in nature. Being binary in form, they should be categorical. Let’s convert their classes to factor. This time, we’ll use simple for and if loops.

```{r}
#convert numeric to factor - train
for (f in names(cd_train[, c(14:20)])) {
  if( class(cd_train[, c(14:20)] [[f]]) == "numeric"){
    levels <- unique(cd_train[, c(14:20)][[f]])
    cd_train[, c(14:20)][[f]] <- as.factor(factor(cd_train[, c(14:20)][[f]], levels = levels))
    }
}

#convert numeric to factor - test
for (f in names(cd_test[, c(13:18)])) {
  if( class(cd_test[, c(13:18)] [[f]]) == "numeric"){
    levels <- unique(cd_test[, c(13:18)][[f]])
    cd_test[, c(13:18)][[f]] <- as.factor(factor(cd_test[, c(13:18)][[f]], levels = levels))
    }
}
```

These loops say – ‘for every column name which falls column number 14 to 20 of cd_train / cd_test data frame, if the class of those variables in numeric, take out the unique value from those columns as levels and convert them into a factor (categorical) variables.

Let’s create some new features now.

```{r}
#Total_Income
cd_train$Total_Income <- cd_train$ApplicantIncome + cd_train$CoapplicantIncome
cd_test$Total_Income <- cd_test$ApplicantIncome + cd_test$CoapplicantIncome

#Income by loan
cd_train$Income_by_loan <- cd_train$Total_Income/cd_train$LoanAmount
cd_test$Income_by_loan <- cd_test$Total_Income/cd_test$LoanAmount

#change variable class
cd_train$Loan_Amount_Term <- as.numeric(cd_train$Loan_Amount_Term)
cd_test$Loan_Amount_Term <- as.numeric(cd_test$Loan_Amount_Term)

#Loan amount by term
cd_train$Loan_amount_by_term <- cd_train$LoanAmount/cd_train$Loan_Amount_Term
cd_test$Loan_amount_by_term <- cd_test$LoanAmount/cd_test$Loan_Amount_Term
```

While creating new features(if they are numeric), we must check their correlation with existing variables as there are high chances often. Let’s see if our new variables too happens to be correlated:

```{r}
#splitting the data based on class
az <- split(names(cd_train), sapply(cd_train, function(x){ class(x)}))

#creating a data frame of numeric variables
xs <- cd_train[az$numeric]

#check correlation
cor(xs)
```

As we see, there exists a very high correlation of Total_Income with ApplicantIncome. It means that the new variable isn’t providing any new information. Thus, this variable is not helpful for modeling data.

Now we can remove the variable.

```{r}
cd_train$Total_Income <- NULL
cd_test$Total_Income <- NULL
```

There is still enough potential left to create new variables. Before proceeding, I want you to think deeper on this problem and try creating newer variables. After doing so much modifications in data, let’s check the data again:

```{r}
summarizeColumns(cd_train)
summarizeColumns(cd_test)
```

# Machine Learning

Until here, we’ve performed all the important transformation steps except normalizing the skewed variables. That will be done after we create the task.

As explained in the beginning, for mlr, a task is nothing but the data set on which a learner learns. Since, it’s a classification problem, we’ll create a classification task. So, the task type solely depends on type of problem at hand.

```{r}
#create a task
trainTask <- makeClassifTask(data = cd_train,target = "Loan_Status")
testTask <- makeClassifTask(data = cd_test, target = "Loan_Status")
```

```{r}
trainTask
```

As you can see, it provides a description of cd_train data. However, an evident problem is that it is considering positive class as N, whereas it should be Y. Let’s modify it:

```{r}
trainTask <- makeClassifTask(data = cd_train,target = "Loan_Status", positive = "Y")
```

For a deeper view, you can check your task data using str(getTaskData(trainTask)).

Now, we will normalize the data. For this step, we’ll use normalizeFeatures function from mlr package. By default, this packages normalizes all the numeric features in the data. Thankfully, only 3 variables which we have to normalize are numeric, rest of the variables have classes other than numeric.

```{r}
#normalize the variables
trainTask <- normalizeFeatures(trainTask,method = "standardize")
testTask <- normalizeFeatures(testTask,method = "standardize")
```

Before we start applying algorithms, we should remove the variables which are not required.

```{r}
trainTask <- dropFeatures(task = trainTask,features = c("Loan_ID","Married.dummy"))
```

MLR package has an in built function which returns the important variables from data. Let’s see which variables are important. Later, we can use this knowledge to subset out input predictors for model improvement. While running this code, R might prompt you to install ‘FSelector’ package, which you should do.

```{r}
#Feature importance
im_feat <- generateFilterValuesData(trainTask, method = c("information.gain","chi.squared"))
plotFilterValues(im_feat,n.show = 20)

#to launch its shiny application
plotFilterValuesGGVIS(im_feat)
```

If you are still wondering about information.gain, let me provide a simple explanation. Information gain is generally used in context with decision trees. Every node split in a decision tree is based on information gain. In general, it tries to find out variables which carries the maximum information using which the target class is easier to predict.

Let’s start modeling now. I won’t explain these algorithms in detail but I’ve provided links to helpful resources. We’ll take up simpler algorithms at first and end this tutorial with the complexed ones.

With MLR, we can choose & set algorithms using makeLearner. This learner will train on trainTask and try to make predictions on testTask.


# Quadratic Discriminant Analysis (QDA)

In general, qda is a parametric algorithm. Parametric means that it makes certain assumptions about data. If the data is actually found to follow the assumptions, such algorithms sometime outperform several non-parametric algorithms. Read More.

```{r}
#load qda 
qda.learner <- makeLearner("classif.qda", predict.type = "response")

#train model
qmodel <- train(qda.learner, trainTask)

#predict on test data
qpredict <- predict(qmodel, testTask)

#create submission file
submit <- data.frame(Loan_ID = test$Loan_ID, Loan_Status = qpredict$data$response)
write.csv(submit, "submit1.csv",row.names = F)
```

Upload this submission file and check your leaderboard rank (wouldn’t be good). Our accuracy is ~ 71.5%. I understand, this submission might not put you among the top on leaderboard, but there’s along way to go. So, let’s proceed.

# Logistic Regression

This time, let’s also check cross validation accuracy. Higher CV accuracy determines that our model does not suffer from high variance and generalizes well on unseen data.

```{r}
#logistic regression
logistic.learner <- makeLearner("classif.logreg",predict.type = "response")

#cross validation (cv) accuracy
cv.logistic <- crossval(learner = logistic.learner,task = trainTask,iters = 3,stratify = TRUE,measures = acc,show.info = F)
```

Similarly, you can perform CV for any learner. Isn’t it incredibly easy? So, I’ve used stratified sampling with 3 fold CV. I’d always recommend you to use stratified sampling in classification problems since it maintains the proportion of target class in n folds. We can check CV accuracy by:

```{r}
cv.logistic$aggr

cv.logistic$measures.test
```

Now, we’ll train the model and check the prediction accuracy on test data.

```{r}
#train model
fmodel <- train(logistic.learner,trainTask)
getLearnerModel(fmodel)

#predict on test data
fpmodel <- predict(fmodel, testTask)

#create submission file
submit <- data.frame(Loan_ID = test$Loan_ID, Loan_Status = fpmodel$data$response)
write.csv(submit, "submit2.csv",row.names = F)
```

Woah! This algorithm gave us a significant boost in accuracy. Moreover, this is a stable model since our CV score and leaderboard score matches closely. This submission returns accuracy of 79.16%. Good, we are improving now. Let’s get ahead to the next algorithm.

# Decision Tree

A decision tree is said to capture non-linear relations better than a logistic regression model. Let’s see if we can improve our model further. This time we’ll hyper tune the tree parameters to achieve optimal results. To get the list of parameters for any algorithm, simply write (in this case rpart):

```{r}
getParamSet("classif.rpart")
```

This will return a long list of tunable and non-tunable parameters. Let’s build a decision tree now. Make sure you have installed the rpart package before creating the tree learner:

```{r}
makeatree <- makeLearner("classif.rpart", predict.type = "response")
```


```{r}
set_cv <- makeResampleDesc("CV",iters = 3L)
```

I’m doing a 3 fold CV because we have less data. Now, let’s set tunable parameters:

```{r}
#Search for hyperparameters
gs <- makeParamSet(
  makeIntegerParam("minsplit",lower = 10, upper = 50),
  makeIntegerParam("minbucket", lower = 5, upper = 50),
  makeNumericParam("cp", lower = 0.001, upper = 0.2)
)
```

As you can see, I’ve set 3 parameters. minsplit represents the minimum number of observation in a node for a split to take place. minbucket says the minimum number of observation I should keep in terminal nodes. cp is the complexity parameter. The lesser it is, the tree will learn more specific relations in the data which might result in overfitting.

```{r}
#do a grid search
gscontrol <- makeTuneControlGrid()

#hypertune the parameters
stune <- tuneParams(learner = makeatree, resampling = set_cv, task = trainTask, par.set = gs, control = gscontrol, measures = acc)
```

You may go and take a walk until the parameter tuning completes. May be, go catch some pokemons! It took 15 minutes to run at my machine. I’ve 8GB intel i5 processor windows machine.

```{r}
#check best parameter
stune$x
```

It returns a list of best parameters. You can check the CV accuracy with:

```{r}
#cross validation result
stune$y
```

Using setHyperPars function, we can directly set the best parameters as modeling parameters in the algorithm.

```{r}
#using hyperparameters for modeling
t.tree <- setHyperPars(makeatree, par.vals = stune$x)

#train the model
t.rpart <- train(t.tree, trainTask)
getLearnerModel(t.rpart)

#make predictions
tpmodel <- predict(t.rpart, testTask)

#create a submission file
submit <- data.frame(Loan_ID = test$Loan_ID, Loan_Status = tpmodel$data$response)
write.csv(submit, "submit3.csv",row.names = F)


```

Decision Tree is doing no better than logistic regression. This algorithm has returned the same accuracy of 79.14% as of logistic regression. So, one tree isn’t enough. Let’s build a forest now.

# Random Forest

Random Forest is a powerful algorithm known to produce astonishing results. Actually, it’s prediction derive from an ensemble of trees. It averages the prediction given by each tree and produces a generalized result. From here, most of the steps would be similar to followed above, but this time I’ve done random search instead of grid search for parameter tuning, because it’s faster.

```{r}
getParamSet("classif.randomForest")

rf <- makeLearner("classif.randomForest", predict.type = "response", par.vals = list(ntree = 200, mtry = 3))
rf$par.vals <- list(importance = TRUE)

#set tunable parameters
#grid search to find hyperparameters
rf_param <- makeParamSet(
  makeIntegerParam("ntree",lower = 50, upper = 500),
  makeIntegerParam("mtry", lower = 3, upper = 10),
  makeIntegerParam("nodesize", lower = 10, upper = 50)
)

#let's do random search for 50 iterations
rancontrol <- makeTuneControlRandom(maxit = 50L)
```
Though, random search is faster than grid search, but sometimes it turns out to be less efficient. In grid search, the algorithm tunes over every possible combination of parameters provided. In a random search, we specify the number of iterations and it randomly passes over the parameter combinations. In this process, it might miss out some important combination of parameters which could have returned maximum accuracy, who knows.

```{r}
#set 3 fold cross validation
set_cv <- makeResampleDesc("CV",iters = 3L)

#hypertuning
rf_tune <- tuneParams(learner = rf, resampling = set_cv, task = trainTask, par.set = rf_param, control = rancontrol, measures = acc)
```

Now, we have the final parameters. Let’s check the list of parameters and CV accuracy.

```{r}
#cv accuracy
rf_tune$y
acc.test.mean 
0.8192571

#best parameters
rf_tune$x

```

Let’s build the random forest model now and check its accuracy.

```{r}
#using hyperparameters for modeling
rf.tree <- setHyperPars(rf, par.vals = rf_tune$x)

#train a model
rforest <- train(rf.tree, trainTask)
getLearnerModel(t.rpart)

#make predictions
rfmodel <- predict(rforest, testTask)

#submission file
submit <- data.frame(Loan_ID = test$Loan_ID, Loan_Status = rfmodel$data$response)
write.csv(submit, "submit4.csv",row.names = F)

```

No new story to cheer about. This model too returned an accuracy of 79.14%. So, try using grid search instead of random search, and tell me in comments if your model improved.


# SVM
Support Vector Machines (SVM) is also a supervised learning algorithm used for regression and classification problems. In general, it creates a hyperplane in n dimensional space to classify the data based on target class. Let’s step away from tree algorithms for a while and see if this algorithm can bring us some improvement.

Since, most of the steps would be similar as performed above, I don’t think understanding these codes for you would be a challenge anymore.

```{r}
#load svm
getParamSet("classif.ksvm") #do install kernlab package 
ksvm <- makeLearner("classif.ksvm", predict.type = "response")

#Set parameters
pssvm <- makeParamSet(
  makeDiscreteParam("C", values = 2^c(-8,-4,-2,0)), #cost parameters
  makeDiscreteParam("sigma", values = 2^c(-8,-4,0,4)) #RBF Kernel Parameter
)

#specify search function
ctrl <- makeTuneControlGrid()

#tune model
res <- tuneParams(ksvm, task = trainTask, resampling = set_cv, par.set = pssvm, control = ctrl,measures = acc)

#CV accuracy
res$y

#set the model with best params
t.svm <- setHyperPars(ksvm, par.vals = res$x)

#train
par.svm <- train(ksvm, trainTask)

#test
predict.svm <- predict(par.svm, testTask)

#submission file
submit <- data.frame(Loan_ID = test$Loan_ID, Loan_Status = predict.svm$data$response)
write.csv(submit, "submit5.csv",row.names = F)
```


This model returns an accuracy of 77.08%. Not bad, but lesser than our highest score. Don’t feel hopeless here. This is core machine learning. ML doesn’t work unless it gets some good variables. May be, you should think longer on feature engineering aspect, and create more useful variables. Let’s do boosting now.

# GBM

Now you are entering the territory of boosting algorithms. GBM performs sequential modeling i.e after one round of prediction, it checks for incorrect predictions, assigns them relatively more weight and predict them again until they are predicted correctly.

```{r}
#load GBM
getParamSet("classif.gbm")
g.gbm <- makeLearner("classif.gbm", predict.type = "response")

#specify tuning method
rancontrol <- makeTuneControlRandom(maxit = 50L)

#3 fold cross validation
set_cv <- makeResampleDesc("CV",iters = 3L)

#parameters
gbm_par<- makeParamSet(
  makeDiscreteParam("distribution", values = "bernoulli"),
  makeIntegerParam("n.trees", lower = 100, upper = 1000), #number of trees
  makeIntegerParam("interaction.depth", lower = 2, upper = 10), #depth of tree
  makeIntegerParam("n.minobsinnode", lower = 10, upper = 80),
  makeNumericParam("shrinkage",lower = 0.01, upper = 1)
)

```


n.minobsinnode refers to the minimum number of observations in a tree node. shrinkage is the regulation parameter which dictates how fast / slow the algorithm should move.

```{r}
#tune parameters
tune_gbm <- tuneParams(learner = g.gbm, 
                       task = trainTask,
                       resampling = set_cv,
                       measures = acc,
                       par.set = gbm_par,
                       control = rancontrol)

#check CV accuracy
tune_gbm$y

#set parameters
final_gbm <- setHyperPars(learner = g.gbm, par.vals = tune_gbm$x)

#train
to.gbm <- train(final_gbm, traintask)

#test 
pr.gbm <- predict(to.gbm, testTask)

#submission file
submit <- data.frame(Loan_ID = test$Loan_ID, Loan_Status = pr.gbm$data$response)
write.csv(submit, "submit6.csv",row.names = F)
```

The accuracy of this model is 78.47%. GBM performed better than SVM, but couldn’t exceed random forest’s accuracy. Finally, let’s test XGboost also.

 

# Xgboost
Xgboost is considered to be better than GBM because of its inbuilt properties including first and second order gradient, parallel processing and ability to prune trees. General implementation of xgboost requires you to convert the data into a matrix. With mlr, that is not required.

As I said in the beginning, a benefit of using this (MLR) package is that you can follow same set of commands for implementing different algorithms.

```{r}
#load xgboost
set.seed(1001)
getParamSet("classif.xgboost")

#make learner with inital parameters
xg_set <- makeLearner("classif.xgboost", predict.type = "response")
xg_set$par.vals <- list(
  objective = "binary:logistic",
  eval_metric = "error",
  nrounds = 250
)

#define parameters for tuning
xg_ps <- makeParamSet(
  makeIntegerParam("nrounds",lower=200,upper=600),
  makeIntegerParam("max_depth",lower=3,upper=20),
  makeNumericParam("lambda",lower=0.55,upper=0.60),
  makeNumericParam("eta", lower = 0.001, upper = 0.5),
  makeNumericParam("subsample", lower = 0.10, upper = 0.80),
  makeNumericParam("min_child_weight",lower=1,upper=5),
  makeNumericParam("colsample_bytree",lower = 0.2,upper = 0.8)
)

#define search function
rancontrol <- makeTuneControlRandom(maxit = 100L) #do 100 iterations

#3 fold cross validation
set_cv <- makeResampleDesc("CV",iters = 3L)

#tune parameters
xg_tune <- tuneParams(learner = xg_set, task = trainTask, resampling = set_cv,measures = acc,par.set = xg_ps, control = rancontrol)

#set parameters
xg_new <- setHyperPars(learner = xg_set, par.vals = xg_tune$x)

#train model
xgmodel <- train(xg_new, trainTask)

#test model
predict.xg <- predict(xgmodel, testTask)

#submission file
submit <- data.frame(Loan_ID = test$Loan_ID, Loan_Status = predict.xg$data$response)
write.csv(submit, "submit7.csv",row.names = F)
```


Terrible XGBoost. This model returns an accuracy of 68.5%, even lower than qda. What could happen ? Overfitting. So, this model returned CV accuracy of ~ 80% but leaderboard score declined drastically, because the model couldn’t predict correctly on unseen data.

 
What can you do next? Feature Selection ?
For improvement, let’s do this. Until here, we’ve used trainTask for model building. Let’s use the knowledge of important variables. Take first 6 important variables and train the models on them. You can expect some improvement. To create a task selecting important variables, do this:

```{r}
#selecting top 6 important features
top_task <- filterFeatures(trainTask, method = "rf.importance", abs = 6)
```

So, I’ve asked this function to get me top 6 important features using the random forest importance feature. Now, replace top_task with trainTask in models above, and tell me in comments if you got any improvement.

Also, try to create more features. The current leaderboard winner is at ~81% accuracy. If you have followed me till here, don’t give up now.


# End Notes
The motive of this article was to get you started with machine learning techniques. These techniques are commonly used in industry today. Hence, make sure you understand them well. Don’t use these algorithms as black box approaches, understand them well. I’ve provided link to resources.

What happened above, happens a lot in real life. You’d try many algorithms but wouldn’t get improvement in accuracy. But, you shouldn’t give up. Being a beginner, you should try exploring other ways to achieve accuracy. Remember, no matter how many wrong attempts you make, you just have to be right once.

You might have to install packages while loading these models, but that’s one time only. If you followed this article completely, you are ready to build models. All you have to do is, learn the theory behind them.

Did you find this article helpful ? Did you try the improvement methods I listed above ? Which algorithm gave you the max. accuracy? Share your observations / experience in the comments below.
